---
title: "HW3"
author: "Shayne Estill (ske2118)"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r, include = FALSE}
library(p8105.datasets)
data("ny_noaa")
```

The "ny_noaa" dataset has `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. From the NY NOAA information page, The resulting R dataset ny_noaa contains variables:

id: Weather station ID
date: Date of observation
prcp: Precipitation (tenths of mm)
snow: Snowfall (mm)
snwd: Snow depth (mm)
tmax: Maximum temperature (tenths of degrees C)
tmin: Minimum temperature (tenths of degrees C)

The variable types include id(chr), date(date), prcp(int), snow(int), snwd(int), tmax(chr), and tmin(chr). 

There is a significant amount of missing data. There are 
`r sum(is.na(ny_noaa$prcp))` missing precipitation values, 
`r sum(is.na(ny_noaa$snow))` missing snow values, 
`r sum(is.na(ny_noaa$snwd))` missing snow depth values, 
`r sum(is.na(ny_noaa$tmax))` missing max temperature values, 
`r sum(is.na(ny_noaa$tmin))` missing min temperature values, 
`r sum(is.na(ny_noaa$date))` missing date values,
`r sum(is.na(ny_noaa$id))` missing ID values 



#Do some data cleaning. Create separate variables for year, month, and day. 

```{r, include = FALSE}
ny_noaa_cleaned <- ny_noaa |>
    janitor::clean_names() |>
    mutate(
    year = lubridate::year(date),
    month = lubridate::month(date),
    day = lubridate::day(date)) |>
    mutate(
      year = as.character(year),
      month = as.character(month),
      day = as.character(day)
    )
```

I also changed the new variables for year, month, and day to character. 


###Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

The max tmax value is `r max(ny_noaa$tmax, na.rm = TRUE)` . 
The min tmin value is `r min(ny_noaa$tmin, na.rm = TRUE)`.
This seems unreasonable for Celsius measures. 

The max precipitation value is `r max(ny_noaa$prcp, na.rm = TRUE)`.
This seems unreasonable for tenths of a mm. 

The max snowfall value is `r max(ny_noaa$snow, na.rm = TRUE)`.
These values don't make sense. We should convert these variables from
character and int to double. 

```{r}
ny_noaa_cleaned <-ny_noaa_cleaned |>
  mutate(
    tmax = as.double(tmax, digits = 2),
    tmin = as.double(tmin, digits = 2),
    snow = as.double(snow, digits = 2),
    snwd = as.double(snow, digits = 2)
  )
```

```{r}
ny_noaa_cleaned <- ny_noaa_cleaned |>
  mutate(
    tmax = tmax / 10,
    tmin = tmin / 10
  )
```

The max precipitation value is `r max(ny_noaa_cleaned$prcp, na.rm = TRUE)`
The max tmax value is `r max(ny_noaa_cleaned$tmax, na.rm = TRUE)`
The min tmin value is `r min(ny_noaa_cleaned$tmin, na.rm = TRUE)`

For snowfall, what are the most commonly observed values? Why?

```{r}
ny_noaa_cleaned |>
  ggplot(aes(x = snow)) + 
  geom_histogram() 

ny_noaa_cleaned |> 
  group_by(snow) |> 
  summarize(count = n()) |> 
  arrange(desc(count))
```

From the histogram and output, the most commonly observed snowfall values are zero, NA, 25 and 13. This is because the majority of the year, there is no snowfall in New York. 


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

The min year is `r min(ny_noaa_cleaned$year, na.rm = TRUE)`.
The max year is `r max(ny_noaa_cleaned$year, na.rm = TRUE)`.


```{r}
avg_maxtemp <- ny_noaa_cleaned |>
  filter(month == 1| month == 7 ) |>
  group_by(id, month, year) |>
  summarize(avg_max_temp = mean(tmax, na.rm = TRUE))

avg_maxtemp |>
  ggplot(aes(x = year, y = avg_max_temp)) + 
  geom_point(aes(color = month)) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ month)
```

The average max temperature in January is much lower than the average max
temperature in July across all years. There does not appear to be too much of an outlier. 


Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option). Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
tmax_tmin = 
  ny_noaa_cleaned |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() 

snowfall_dist = 
  ny_noaa_cleaned |>
  filter(snow < 100, snow > 0) |>
  ggplot(aes(x = snow, y = as.factor(year))) + 
  geom_density_ridges(scale = 0.5)

tmax_tmin + snowfall_dist
```

###Problem 2###

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}
demographic_df = read_csv(file = "./HW3_2_files/nhanes_covar.csv",   
  skip = 4) |>
janitor::clean_names() |>
    filter(age > 20) |>
    drop_na() |>
    mutate(
      seqn = factor(seqn),
      sex = factor(sex),
      education = factor(education)
    )

```

```{r}
accel_df = read_csv(file = "./HW3_2_files/nhanes_accel.csv") |>
  janitor::clean_names() |>
  drop_na() |>
  mutate(
    seqn = factor(seqn)
  )

```

```{r}
accel_tidy_df = 
  pivot_longer(
    accel_df,
    min1:min1440,
    names_to = "obs",
    values_to = "min"
  )
  
merged_df = 
  left_join(
    accel_tidy_df,
    demographic_df,
    by = NULL,
    copy = FALSE
  )
```


Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
counts_df <-
merged_df |>
  group_by(education, sex) |>
  count()

pivot_wider(
  counts_df, 
  names_from = "sex", 
  values_from = n)

knitr::kable(counts_df)
```


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}

```


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}

```



###Problem 3###